### 2022-10-27 Readme.Rmd
rewamp, pcm

---
title: Identifying covariate traits of trance and possession phenomena in the Ethnographic
  Atlas using boosting and nested regression
author: "Péter Rácz"
date: "`r format(Sys.Date(),'%e %B, %Y')`"
output:
  md_document:
    variant: markdown_github
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.path = 'figures/', fig.width = 8, fig.height = 8)
```

```{r header}
setwd('~/Github/Racz2023trance/')

library(tidyverse)
library(glue)
library(knitr)
library(patchwork)
library(corrplot)

ld = read_tsv('data/dat_long.tsv')
wd = read_tsv('data/dat_wide.tsv')
bf = read_tsv('models/bayes_factors.tsv')
cm1a = read_tsv('models/conf_matrix11model.tsv')
cm1b = read_tsv('models/conf_matrix11test.tsv')
varimp1 = read_tsv('models/varimp11.tsv')
sccs = filter(ld, in_sccs)
```

We want to look at three questions. What sort of social factors have been proposed to co-occur with trance and possession phenomena? Using the Ethnographic Atlas and the Standard Cross-Cultural Sample, can we get more confident that these factors do co-occur with such phenomena across cultures? Using phylogenetic comparative methods, can we make claims on the causative structure between the co-occurent factors?

## 1. Trance and possession phenomena have societal covariates in the literature

Many people have argued that trance induced by possession is not a random, idiosyncratic pathology but a recurrent psycho-social trait that is at least partly motivated by other social factors.

High overall social complexity and social rigidity -- the resulting power dynamics -- might translate to possession trance practices. Such practices might serve to express or relay perceived social inflexibility and inequality. This is echoed in the medical anthropology literature (@during2011critical,@dein2012religion) and ties in with broader questions of inequality and mental health (@nguyen2003anthropology). Various causative models have been proposed to account for the emergence of possession trance and, specifically, shamanism in a social setting (@winkelman2015shamanism, @singh2018cultural, @wood2018trance).

Some or most groups can feel stuck in a society. If this makes trance and possession more likely, we should see correlations between measures of social complexity, organisation, and inequality and possession and trance in cross-cultural data. Many such correlations have been proposed by previous work. Here is a list, cross-referenced with @wood2018complexity:

| Relationship                                                                             | Reference                |
|:-----------------------------------|:-----------------------------------|
| Social complexity, rigidity, gender role differentiation                                    | @bourguignon1968         |
| Women's social participation                                                             | @bourguignon1973religion |
| Slavery, role differentiation, and structure differentiation | @greenbaum1973           |
| Increased social participation                                | @swanson1978             |
| Social complexity                                     | @winkelman1986           |
| Social flexibility, cognatic kinship                              | @shaara1992              |
| Decreased social participation                                | @douglas2004natural      |
| Social rigidity, monotheism                                                                               | @wood2018complexity      |

Several works made the same point or emphasised various aspects of the same relationship. The above table is somewhat reductive. The argument is spelt out in the main text.

## 2. This can be mapped to the Ethnographic Atlas

The use of cross-cultural typological data in general and the Ethnographic Atlas in particular has precedents. @bourguignon1968, @bourguignon1973religion, and @greenbaum1973 use the Ethnographic Atlas to look for correlates of trance and possession trance. @wood2018complexity, @shaara1992, @winkelman1986, and @douglas2004natural build on this research by specifying and compiling their own data sets. The present paper uses the Ethnographic Atlas as made available in the D-Place database, @kirby2016d.

The Atlas records the presence of trance and possession phenomena in a given society, using the following variable (EA112):

```{r }
trance = sccs %>% 
  filter(var_id == 'EA112') %>% 
  count(var_name,code,var_description) %>% 
  rename(name = var_name, description = var_description, "n in SCCS" = n) %>% 
  mutate(
    possession_trance = case_when(
      code %in% c(3,4,6,7) ~ 'possession trance',
      code %in% c(1,2,8,5) ~ 'other'
    ) %>% 
      fct_relevel('possession trance')
  )

ld %>% 
  filter(var_title == 'Trance states', !is.na(code)) %>% 
  count(code, var_description, name = 'n in EA') %>% 
  left_join(trance) %>% 
  arrange(code) %>% 
  rename(Coding = code, Description = var_description) %>% 
  select(Coding,Description,possession_trance,`n in EA`,`n in SCCS`) %>% 
  kable('simple')
```

This is basically a truth table of whether possession beliefs or trance states are present in a given society, with some causal links between the two (if possession happens it leads to trance, if trance happens it's because of possession, etc). We can think of different groupings (trance phenomena encompass levels 1,3,4,5,6,7, possession phenomena encompass 2,3,4,5,6,7, specifically possession trance covers 3,4,6,7 but not 1,2,5,8).

The Atlas lists a set of variables that tie in with the observed covariates in the literature. These are either taken literally from @bourguignon1968, @bourguignon1973religion, and @greenbaum1973 or adapt variables used by @wood2018complexity, @shaara1992, @winkelman1986, and @douglas2004natural.

```{r }
ld %>% 
  filter(var_title != 'Trance states') %>% 
  distinct(var_id,var_title,var_definition) %>%
  arrange(var_id) %>% 
  rename(ID = var_id, Name = var_title, Definition = var_definition) %>% 
  kable('simple')
```

## 3. How do we look for correlations between trance and possession and their proposed societal covariates?

There are eight types of trance and possession phenomena recorded in the Atlas and a wide set of possible covariates, all with a number of levels (see the Appendix).

The data are challenging in two ways. The variables are themselves correlated, either because they measure different facets of the same thing (like how complex a society is; see the figure below) or because they were adapted together by socially or geographically close societies (Galton's problem). In addition, data are missing from the Atlas in a non-random manner (e.g. if a society is hard to access, it is harder to provide a population size estimate for it). The correlations of the pairwise complete observations of the relevant predictors in the Ethnographic Atlas can be seen below.

```{r }
cord = ld %>% 
  filter(!var_title %in% c('Trance states','Societal rigidity')) %>% 
  select(soc_id,var_title,code) %>%
  pivot_wider(id_cols = soc_id, names_from = var_title, values_from = code) %>% 
  select(-soc_id)

cor(cord,use="pairwise.complete.obs") %>% 
  corrplot()

```

We can account for Galton's problem (geographic and cultural co-variation) by restricting our initial data to societies in the Standard Cross-Cultural Sample.

The remaining co-variation and the non-random patterns of missing data recommend an ensemble learning method. This starts with societies that have information on trance and possession phenomena. It then builds a lot of tiny models, using random samples of the data and random subsets of the covariates. Then, we compare the tiny models with each other to assess the overall importance of our covariates. Building tiny models in parallel is called bagging. This is how a random forest works, growing many classification or regression trees and aggregating over the results. One step further is to train models on the mistakes of the previous models. This is called [boosting](https://www.ibm.com/cloud/learn/boosting) and usually results in higher training and prediction accuracy.

## 4. Building a boosting model

In this analysis, we will use a gradient boosting model on the data from the Ethnographic Atlas, subsetting it to only include societies in the Standard Cross-Cultural Sample.

Gradient boosting is an ensemble learning method which uses decision trees. A decision tree is a directed tree graph in which each node represents a decision or split based on a predictor variable and each leaf is an individual data point. It can be used to classify leaves by grouping them under different node structures. A gradient boosting algorithm fits such trees in a sequence, training each on the errors of the previous one, minimising overall error. It stops when further iterations do not result in meaningful improvement on model accuracy (according to pre-set criteria) [@cook2016practical,@rhys2020machine,@smith2017decision].

We only fit the model on societies which (a) have information on whether trance and possession phenomena are present and (b) come from the standard cross-cultural sample.

Following @bourguignon1968, we reduce the complexity of the outcome variable to two levels: whether a society has *possession trance* or not. The predictors are the possible covariates identified above. Some predictors, like `EA032` (Jurisdictional hierarchy of local community) or `EA113` can be construed to follow a scale. These are input as *ordered factors*. Others, like `EA053` (Sex differences in animal husbandry) or `EA043` (Major type of descent) can be carved up into categories: These are *one-hot encoded*. For details, see the Appendix.

We use *hyperparameter tuning* to find the best possible model.  The model is fit on a training set of `length(unique(sccs$soc_id))` societies with information on trance and possession in the standard cross-cultural sample. The hyperparameter grid is printed in the Appendix.

## 5. Results on the training sample

Looking back at the table in the previous section, we see that only about 10% of societies have no trance or possession phenomena. From the rest, a majority have possession trance in some form. Following @bourguignon1968 we can reduce this to a binary: 81 societies have possession trance, 70 have other forms of trance and possessior or show neither.

```{r }
trance %>% 
  group_by(possession_trance) %>% 
  summarise(count = sum(`n in SCCS`)) %>% 
  kable('simple')
```

We compare training information on each society with how our best model categorises each society. This is our model's confusion matrix. `0` is absence of possession trance, `1` is presence of possession trance. The rows add up to values in the original, the columns are the model's predictions. The last two columns show the error rate. For example, there are 71 societies in the Standard Cross-Cultural Sample that have possession trance, out of these, the model thinks 64 have possession trance and 7 do not (second row). This is an error rate of 7/71=.1.

```{r }

cm1a %>% 
  mutate(original = c(names(cm1a)[!names(cm1a) %in% c('Error','Rate')],'Total')) %>% 
  relocate(original) %>% 
  kable('simple', digits = 2)

# precision = true 1/tru1 + false 1
# recall = true 1/true1 + false0
precision = 64 / (64+12)
recall = 64 / (64+7)
f = (2 * precision * recall) / (precision + recall)
# chisq.test(c(c(68,12),c(7,64)))
```

The model has an F-measure of `r round(f,2)`, with a penalty on slightly overgeneralising possession trance. Though such comparisons are hard to make, this is not particularly accurate for a binary categorisation problem. This is not surprising -- it is unlikely that any complex cultural phenomenon could be reduced to a small number of cross-cultural predictors.

### 6. Results on the EA data

We can ask the model to make predictions for the entire Ethnographic Atlas (where trance and possession data are available) -- including the Standard Cross-Cultural Sample subset. Generally, you want to separate the training and test sets in Machine Learning and not do this, but the 'everything except the Standard Cross-Cultural Sample' subset of the Atlas would be an odd test set.

Here is the confusion matrix for all `r length(unique(ld$soc_id))` societies in the Atlas where trance and possession information are available:

```{r }

cm1b %>% 
  kable('simple')

# precision = true 1/true1+false 1
# recall = true 1/true1 + false0
precision2 = 194 / (194+113)
recall2 = 194 / (194+87)
f2 = (2 * precision2 * recall2) / (precision2 + recall2)
chisq.test(c(c(264,113),c(87,194)))
```

The F-measue of this model is `r round(f2,2)`. Accuracy here is much worse than for the training data only. That is partly due to model bias and at least partly because the model is ignorant of the phylogenetic signal which might result in patterns that are unexpected from a purely correlational point of view. However, the model is still more accurate than it would be by chance.

We now take a look at the variables that were most important in the model itself in predicting the presence of possession trance in a given society, with a cutoff of 10/47.

```{r }
vars = ld %>% 
  filter(var_title != 'Trance states') %>% 
  distinct(var_id,var_title,var_definition,var_name) %>%
  arrange(var_id) %>% 
  group_by(var_id, var_title, var_definition) %>% 
  nest() %>% 
  mutate(levels = map(data, ~ paste(.$var_name, collapse = ', '))) %>% 
  select(var_id, var_title, levels) %>% 
  unnest(cols = c(levels)) %>% 
  rename(ID = var_id, Name = var_title)

vars2 = varimp1 %>% 
  slice(1:10) %>% 
  mutate(ID = str_extract(variable, '^EA...')) %>% 
  left_join(vars, by = "ID") %>% 
  select(ID, Name, relative_importance)

vars2 %>% kable(digits = 2, 'simple')
```

We now look at the direction of these effects (does e.g.\ increased jurisdictional hierarchy go with more or less prevalent possession trance?).

The figure below shows the proportion of societies with possession trance (dark blue) and without (light blue) across the ordered levels of the predictors. These are the predictors that the model thinks are important. They come in order of how important they are, according to the model; prevalence of slavery, then family organisation, and so on. The predictor levels are ordered according to coding and generally go from a pattern being less complex or prevalent (top) to more complex or prevalent (bottom). So, for domestic organisation (top right) we go from small nuclear to large extended families. For slavery (top left) we go from absence of slavery to hereditary slavery.

We see a clear correlation between the presence of possession trance and the variable scaling up. E.g. societies in which slavery is more prevalent are more likely to have possession trance than those where it is less prevalent. Societies with more complex extended families are more likely to have possession trance. The across-level distributions are also interesting, e.g. the main distinction for patterns of slavery is for societies without slavery (first factor level) and all others with some form of slavery (subsequent factor levels).

```{r fig.width=10,fig.height=15}
d_out = ld %>% 
  filter(var_id == 'EA112') %>% 
  mutate(
    possession_trance_present = case_when(
      code %in% c(3,4,6,7) ~ 'present',
      code %in% c(1,2,5,8) ~ 'absent'
        )
      ) %>% 
  select(soc_id,possession_trance_present)

d_in = ld %>%
  filter(var_title %in% vars2$Name, !is.na(code)) %>% 
  mutate(
    predictor = fct_reorder(var_name, -as.double(recoded_value)),
    var_title = fct_relevel(var_title, vars2$Name)
    ) %>% 
  select(soc_id,var_definition,var_title,predictor)

d_vis = inner_join(d_in,d_out, by = 'soc_id')

plots = d_vis %>%
  count(var_title,predictor,possession_trance_present) %>% 
  # pivot_wider(names_from = possession_trance_present, values_from = n, values_fill = 0) %>% 
  # mutate(log_odds_pt = log((`TRUE` + 1) / (`FALSE` + 1)) ) %>% 
  group_by(var_title) %>% 
  nest() %>% 
  mutate(
    plot = map2(var_title, data, ~ ggplot(.y, aes(predictor,n,fill = possession_trance_present)) +
                  geom_col(position = 'dodge') +
                  theme_bw() +
                  theme(axis.title=element_blank()) +
                  guides(fill = 'none') +
                  scale_fill_brewer(palette = "Paired") +
                  ggtitle(str_wrap(.x, width = 25)) +
                  coord_flip()
                 )
  ) %>% 
  pull(plot)

wrap_plots(plots, ncol = 2) + plot_annotation(caption = 'Dark blue: possession trance present. Light blue: absent.')
```

## 7. Did these traits develop independently?

under construction.

slavery. domestic organisation.
binary.
austronesian (eeh http://lir.byuh.edu/index.php/pacific/article/view/141)
afro-asiatic (sub-saharan, greenbaum)
atlantic-congo (hamito-semitic! lewis)

bayestraits

The independent model assumes the two traits evolve independently, e.g. the transition
from 0 → 1 in the first trait is independent of the state of the second trait. 
dependent: the traits are correlated and the rate of change in one
trait is dependent on the state of the other
100 stones and 1000

mcmc exp 10 priors
iterations per stone to estimate the marginal likelihood

bayes factor
```{r }
kable(bf, 'simple', digits = 2)
```

lot of forking paths. slavery in atlantic-congo noteworthy. maybe. overall not really.

## 8. Summary

These results go some way in confirming hypotheses on the cross-cultural correlates of possession trance in the literature. @bourguignon1968 explicitly tests the variables of slavery, jurisdictional hierarchy, settlement patterns, and kin groups. @shaara1992 point to a high value placed on female virginity at marriage, captured here by norms of premarital sexual behaviour for girls, @wood2018complexity point to the role of moralising high gods. These are all important variables in our model. @greenbaum1973 and @shaara1992 point to sexual division of labour but this is not relevant in our model.

Note that @shaara1992 use cognatic kinship as a proxy of social complexity (which in itself is robustly supported by our model) while the measure referenced by @wood2018complexity, the presence of moralising high gods, has in itself been argued to correlate with social complexity (see @botero2014ecology and @watts2015broad).

This brings us back to the original predictions in the table above.

| Relationship|Reference|Our model|
|:-|:-|:-|
|less complexity$\rightarrow{}$trance w/o possession|@bourguignon1968|the same measures of complexity are relevant|
|women's social participation|@bourguignon1973religion|labour divisions are not relevant|
|increased role differentiation and slavery practices$\rightarrow$trance and possession|@greenbaum1973|slavery, social and class complexity are relevant|
|main mode of subsistence|@bourgouignon1977|not relevant|
|increased social participation$\rightarrow{}$possession|@swanson1978|indirectly, through domestic and social complexity|
|more complexity$\rightarrow{}$trance w/ possession|@winkelman1986|yes, see above|
|cognatic kinship$\rightarrow{}$more possession and trance|@shaara1992|not directly|
|decreased social participation$\rightarrow{}$possession|@douglas2004natural|yes, see above|
|monotheism|@wood2018complexity|yes|

We can predict some types of trance and possession phenomena in the SCCS subset of the EA. Not all of them. Some of the important co-variates are the ones mentioned by earlier work and the direction of the relationships corroborates the literature.

something something phylogenetic comparative methods

Human society is complicated and cross-cultural data have very low resolution. In this sense, we'd be very suspect of a model that would be too accurate of predicting possession trance from societal covariates. It cannot replace fieldwork and more complex explanatory models either. What it can do is provide an extra leg for such explanatory work, which is broadly the recognised ambition of cross-cultural comparative work in evolutionary anthropology.

Our approach is statistically strict, so the fact that it still shows some accuracy is definitely promising for the validity of the connections drawn by previous work on trance and possession phenomena in particular and large cross-cultural comparisons in general.

## Appendix

### Relevant predictors in the Ethnographic Atlas

A survey of the literature resulted in the following predictors of trance and possession phenomena.

```{r }
kable(vars, 'simple')
```

### Recoded values in the Ethnographic Atlas

The recoded values were used in the model for each predictor. Where the recoded values are numeric, the predictor was used as an ordered factor. Where the recoded values are strings, the predictor was one-hot encoded (such that predictor A with levels 1, 2, and 3 was turned into boolean predictors A==2 and A==3).

```{r }
ld %>% 
  filter(var_id != 'EA112') %>% 
  distinct(var_id,var_title,var_description,recoded_value) %>% 
  arrange(var_id,recoded_value) %>% 
  kable('simple')
```

### Hyperparameter values for the Gradient Boosting Model

```{r, echo=T}
gbm_params1 = list(
  learn_rate = c(0.01,0.02,0.03,0.1),
  max_depth = c(1,3,5,7),
  sample_rate = c(0.4,0.7,0.9,0.95),
  col_sample_rate = c(0.1,0.2,0.3),
  min_split_improvement = c(1e-3,1e-5)
)
```

## References